---
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    toc: yes
---
### 1. Loading and preprocessing the data

#### Set the 'Working directory'
```{r Set working directory}
path <- getwd()
```

#### Read the 'Activity' data
```{r}
path <- getwd()
dt_Activity <- read.table(file.path(path, "repdata-data-activity", "activity.csv"),header = TRUE, sep = ",")
```

#### Convert the 'Interval' value into 'min' and create a new column by combining both 'Date' and 'min' values
```{r}
time <- formatC(dt_Activity$interval/100, 2, format = "f")
dt_Activity$date.time <- as.POSIXct(paste(dt_Activity$date, time), format = "%Y-%m-%d %H.%M", tz = "GMT")
```

### 2. What is mean total number of steps taken per day? 

#### Calculate the total number per day
```{r}
totalSteps <- tapply(dt_Activity$steps, dt_Activity$date, sum, na.rm = TRUE)
```

#### Calculate the mean and median for total steps / day
```{r}
mean(totalSteps)
median(totalSteps)
```

####  Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(totalSteps, xlab = "Total steps", ylab = "Frequency")
```

### 3. What is the average daily activity pattern?

#### Add a 'Time' column
```{r}
dt_Activity$time <- format(dt_Activity$date.time, format = "%H:%M:%S")
dt_Activity$time <- as.POSIXct(dt_Activity$time, format = "%H:%M:%S")
```
#### Calculate the mean steps for every five mins interval
```{r}
meanSteps <- tapply(dt_Activity$steps, dt_Activity$time, mean, na.rm = TRUE)
df_Activity_pattern <- data.frame(Time = as.POSIXct(names(meanSteps)), meanSteps = meanSteps)
```
#### Mean steps vs Time plot
```{r}
library(scales)
ggplot(df_Activity_pattern , aes(Time, meanSteps)) + geom_line() + xlab("Time of day") + 
    ylab("Mean number of steps") + scale_x_datetime(labels = date_format(format = "%H:%M"))
```

### 4. Imputing missing values

#### Calculate the missing value (NA's)
```{r}
summary(dt_Activity$steps)
```

#### Create a new dataset from the original dataset , and fill the 'Missing value'
```{r}
dt_Activity_Impute <- dt_Activity
library(Hmisc)
dt_Activity_Impute$steps <- with(dt_Activity_Impute, impute(steps, mean))
summary(dt_Activity_Impute$steps)
```
#### Compare the mean and median steps for each day between the original data set and the imputed data set

#### Calculate the total number per day from Imputed data set
```{r}
totalSteps_Imputed <- tapply(dt_Activity_Impute$steps, dt_Activity_Impute$date, sum)
```

#### Compare the mean and median - Original vs Imputed data set
```{r}
mean(totalSteps)
median(totalSteps)

mean(totalSteps_Imputed)
median(totalSteps_Imputed)
```

#### Histogram of the total number of steps (Imputed data) taken each day
```{r}
qplot(totalSteps_Imputed, xlab = "Total steps", ylab = "Frequency")
```

### 5. Are there differences in activity patterns between weekdays and weekends?

#### Add a factor column [ a day is 'weekend' or 'weekday']
```{r}
day_type <- function(date){
  if(weekdays(date) %in% c("Saturday", "Sunday")){
      return("weekend")
  }else{
    return("weekday")
  }
}

dayTypes = sapply(dt_Activity_Impute$date.time, day_type)
dt_Activity_Impute$day.type <- as.factor(dayTypes)
```

#### Create a dataframe that stores mean steps value for ' Day type' values
```{r}
meanSteps <- tapply(dt_Activity_Impute$steps, interaction(dt_Activity_Impute$time, dt_Activity_Impute$day.type), mean, na.rm = TRUE)
df_dayType_pattern <- data.frame(Time = as.POSIXct(names(meanSteps)), meanSteps = meanSteps, day.type = as.factor(c("weekend", "weekday")))
```
#### Display two patterns for weekdays and weekends
```{r}
ggplot(df_dayType_pattern , aes(Time, meanSteps)) + geom_line() + xlab("Time of day") + 
    ylab("Mean number of steps") + scale_x_datetime(labels = date_format(format = "%H:%M")) + facet_grid(. ~ day.type)
```
